---
title: "C19 Data and Trends"
output: html_document
---


```{r, echo=FALSE, include=FALSE}

#----------------------------------------------------
# Pulling the coronvirus data from John Hopkins repo
# https://github.com/CSSEGISandData/COVID-19
#----------------------------------------------------
# Setting functions
`%>%` <- magrittr::`%>%`
#----------------------------------------------------
# Pulling confirmed cases

raw_conf <- read.csv(file = "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Confirmed.csv",
                     stringsAsFactors = FALSE)
# Transforming the data from wide to long
# Creating new data frame
df_conf <- raw_conf[, 1:4]

for(i in 5:ncol(raw_conf)){
  print(i)
  raw_conf[,i] <- as.integer(raw_conf[,i])
  raw_conf[,i] <- ifelse(is.na(raw_conf[, i]), 0 , raw_conf[, i])
  
  if(i == 5){
    df_conf[[names(raw_conf)[i]]] <- raw_conf[, i]
  } else {
    df_conf[[names(raw_conf)[i]]] <- raw_conf[, i] - raw_conf[, i - 1]
  }
}


df_conf1 <-  df_conf %>% tidyr::pivot_longer(cols = dplyr::starts_with("X"),
                                             names_to = "date_temp",
                                             values_to = "cases_temp")

# Parsing the date
df_conf1$month <- sub("X", "",
                      strsplit(df_conf1$date_temp, split = "\\.") %>%
                        purrr::map_chr(~.x[1]) )

df_conf1$day <- strsplit(df_conf1$date_temp, split = "\\.") %>%
  purrr::map_chr(~.x[2])


df_conf1$date <- as.Date(paste("2020", df_conf1$month, df_conf1$day, sep = "-"))

# Aggregate the data to daily
df_conf2 <- df_conf1 %>%
  dplyr::group_by(Province.State, Country.Region, Lat, Long, date) %>%
  dplyr::summarise(cases = sum(cases_temp)) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(type = "confirmed",
                Country.Region = trimws(Country.Region),
                Province.State = trimws(Province.State))

head(df_conf2)
tail(df_conf2)
#----------------------------------------------------
# Pulling death cases

raw_death <- read.csv(file = "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Deaths.csv",
                      stringsAsFactors = FALSE)
# Transforming the data from wide to long
# Creating new data frame
df_death <- raw_death[, 1:4]

for(i in 5:ncol(raw_death)){
  print(i)
  raw_death[,i] <- as.integer(raw_death[,i])
  raw_death[,i] <- ifelse(is.na(raw_death[, i]), 0 , raw_death[, i])
  
  if(i == 5){
    df_death[[names(raw_death)[i]]] <- raw_death[, i]
  } else {
    df_death[[names(raw_death)[i]]] <- raw_death[, i] - raw_death[, i - 1]
  }
}


df_death1 <-  df_death %>% tidyr::pivot_longer(cols = dplyr::starts_with("X"),
                                               names_to = "date_temp",
                                               values_to = "cases_temp")

# Parsing the date
df_death1$month <- sub("X", "",
                       strsplit(df_death1$date_temp, split = "\\.") %>%
                         purrr::map_chr(~.x[1]) )

df_death1$day <- strsplit(df_death1$date_temp, split = "\\.") %>%
  purrr::map_chr(~.x[2])


df_death1$date <- as.Date(paste("2020", df_death1$month, df_death1$day, sep = "-"))

# Aggregate the data to daily
df_death2 <- df_death1 %>%
  dplyr::group_by(Province.State, Country.Region, Lat, Long, date) %>%
  dplyr::summarise(cases = sum(cases_temp)) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(type = "death",
                Country.Region = trimws(Country.Region),
                Province.State = trimws(Province.State))

head(df_death2)
tail(df_death2)
#----------------------------------------------------
# Pulling recovered cases

raw_rec <- read.csv(file = "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Recovered.csv",
                    stringsAsFactors = FALSE)
# Transforming the data from wide to long
# Creating new data frame
df_rec <- raw_rec[, 1:4]

for(i in 5:ncol(raw_rec)){
  print(i)
  raw_rec[,i] <- as.integer(raw_rec[,i])
  raw_rec[,i] <- ifelse(is.na(raw_rec[, i]), 0 , raw_rec[, i])
  
  if(i == 5){
    df_rec[[names(raw_rec)[i]]] <- raw_rec[, i]
  } else {
    df_rec[[names(raw_rec)[i]]] <- raw_rec[, i] - raw_rec[, i - 1]
  }
}


df_rec1 <-  df_rec %>% tidyr::pivot_longer(cols = dplyr::starts_with("X"),
                                           names_to = "date_temp",
                                           values_to = "cases_temp")

# Parsing the date
df_rec1$month <- sub("X", "",
                     strsplit(df_rec1$date_temp, split = "\\.") %>%
                       purrr::map_chr(~.x[1]) )

df_rec1$day <- strsplit(df_rec1$date_temp, split = "\\.") %>%
  purrr::map_chr(~.x[2])


df_rec1$date <- as.Date(paste("2020", df_rec1$month, df_rec1$day, sep = "-"))

# Aggregate the data to daily
df_rec2 <- df_rec1 %>%
  dplyr::group_by(Province.State, Country.Region, Lat, Long, date) %>%
  dplyr::summarise(cases = sum(cases_temp)) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(type = "recovered",
                Country.Region = trimws(Country.Region),
                Province.State = trimws(Province.State))

head(df_rec2)
tail(df_rec2)

coronavirus <- dplyr::bind_rows(df_conf2, df_death2, df_rec2) %>%
  dplyr::arrange(date) %>% dplyr::ungroup() %>%
  dplyr::filter(cases != 0)
head(coronavirus)
tail(coronavirus)


rm(list= ls()[! (ls() %in% c('coronavirus'))])








```



```{r, echo=FALSE, include=FALSE}
library(dplyr)
library(stringr)

```


```{r, echo=FALSE, include=FALSE}

inc_countries = c("Mainland China", "Iran", "Italy", "Germany", "US", "Canada", "South Korea", "UK",
                  "France", "Japan")

df = coronavirus %>% filter(Country.Region %in% inc_countries)

#rm(inc_countries)

```

## Worldwide Summary - All dates to current

```{r, echo=FALSE, include=TRUE}

total_cases = coronavirus %>% group_by(type) %>% summarise(cases = sum(cases))
mrate = total_cases[2,2] / total_cases[1,2]
mort_rate = data.frame("mort_rate", mrate)
names(mort_rate) = c("type", "cases")
total_cases = rbind(total_cases, mort_rate)

print(total_cases)

sum = coronavirus %>% group_by(Country.Region) %>% summarise(cases = sum(cases)) %>% arrange(desc(cases))

print(sum)

```

## Worldwide Trends - Timeseries

```{r, echo=FALSE, warning=FALSE}

by_day = coronavirus %>% subset(type == "confirmed") %>% 
  group_by(date) %>%
  summarise(cases = sum(cases))

plot(by_day$date, by_day$cases, type = "b", xlab = "Date", ylab = "Cases", main = "Worldwide Confirmed Cases - Timeseries")


```


## Worldwide Summary - All dates to current (excluding China)

```{r, echo=FALSE, include=TRUE}

total_cases = coronavirus %>% filter(Country.Region != "Mainland China") %>% group_by(type) %>% 
  summarise(cases = sum(cases))
mrate = total_cases[2,2] / total_cases[1,2]
mort_rate = data.frame("mort_rate", mrate)
names(mort_rate) = c("type", "cases")
total_cases = rbind(total_cases, mort_rate)

#add average daily rate of change
del = coronavirus %>% filter(Country.Region != "Mainland China" & date >= '2020-02-16' & type == "confirmed") %>%
  group_by(date) %>% summarise(cases = sum(cases))

del = del %>% mutate(diff = 1 + (cases - lag(cases)  )/ lag(cases)  )




rate_of_change = mean(del$diff[-1])
rate_of_change = data.frame("rate_of_growth", rate_of_change)
names(rate_of_change) = c("type", "cases")
total_cases = rbind(total_cases, rate_of_change)



print(total_cases)


```

## Growth Rate Plot
##### Rate of Growth = 1 + ( (Number of cases per day - Previous day's cases) / Previous day's cases )



```{r, echo=FALSE, warning=FALSE}



plot(del$date, del$diff, type = "b", xlab = "Date", ylab = "Rate of Change", main = "Rate of Change Trend (excluding China) - Timeseries")


```

## Worldwide Trends - Timeseries (excluding China)

```{r, echo=FALSE, warning=FALSE}

by_day = coronavirus %>% subset(type == "confirmed" & Country.Region != "Mainland China") %>% 
  group_by(date) %>%
  summarise(cases = sum(cases))

plot(by_day$date, by_day$cases, type = "b", xlab = "Date", ylab = "Cases", main = "Worldwide Confirmed Cases (excluding China) - Timeseries")


```



## US Totals
```{r, echo=FALSE, warning=FALSE}

us = df %>% subset(Country.Region == "US") %>% group_by(type) %>% summarise(cases = sum(cases))
mrate = us[2,2] / us[1,2]
mort_rate = data.frame("mort_rate", mrate)
names(mort_rate) = c("type", "cases")
us = rbind(us, mort_rate)

print(us)

```


## US Trends - Timeseries

```{r, echo=FALSE, warning=FALSE}

by_day = df %>% subset(Country.Region == "US" & type == "confirmed") %>% 
  group_by(date) %>%
  summarise(cases = sum(cases))

plot(by_day$date, by_day$cases, type = "b", xlab = "Date", ylab = "Cases", main = "US Confirmed Cases - Timeseries")


```



## China Totals
```{r, echo=FALSE, warning=FALSE}

ch = coronavirus %>% subset(Country.Region == "Mainland China") %>% group_by(type) %>% summarise(cases = sum(cases))
mrate = ch[2,2] / ch[1,2]
mort_rate = data.frame("mort_rate", mrate)
names(mort_rate) = c("type", "cases")
ch = rbind(ch, mort_rate)

print(ch)

```


## China Trends - Timeseries - All

```{r, echo=FALSE, warning=FALSE}

by_day = coronavirus %>% subset(Country.Region == "Mainland China" & type == "confirmed") %>% 
  group_by(date) %>%
  summarise(cases = sum(cases))

plot(by_day$date, by_day$cases, type = "b", xlab = "Date", ylab = "Cases", main = "China Confirmed Cases - Timeseries")


```

## China Trends - Timeseries - Feb 14th to current

```{r, echo=FALSE, warning=FALSE}

by_day = coronavirus %>% subset(Country.Region == "Mainland China" & type == "confirmed" & date >= '2020-02-15') %>% 
  group_by(date) %>%
  summarise(cases = sum(cases))

plot(by_day$date, by_day$cases, type = "b", xlab = "Date", ylab = "Cases", main = "China Confirmed Cases - Timeseries")


```


## CA Summary

```{r, echo=FALSE}
us_all = df %>% subset(Country.Region == "US") 
us_all$state = str_sub(us_all$Province.State,-2,-1)
ca = us_all %>% filter(state == "CA")
ca_summary = ca %>% group_by(type) %>% summarise(cases = sum(cases))

mrate = ca_summary[2,2] / ca_summary[1,2]
mort_rate = data.frame("mort_rate", mrate)
names(mort_rate) = c("type", "cases")
ca_summary = rbind(ca_summary, mort_rate)

county = ca %>% filter(type == "confirmed") %>% group_by(Province.State) %>% 
  summarise(confirmed_cases = sum(cases)) %>%
  arrange(desc(confirmed_cases))


print(ca_summary)
print(county)

```


## CA Trend

```{r, echo=FALSE}
ca_trend = ca %>% filter(type == "confirmed") %>% group_by(date)

plot(ca_trend$date, ca_trend$cases, type = "p", xlab = "Date", ylab = "Cases", main = "California Confirmed Cases - Timeseries")

```


## Santa Clara Trend

```{r, echo=FALSE}

sc = ca %>% filter(Province.State == "Santa Clara County, CA" & type == "confirmed") 

plot(sc$date, sc$cases, type = "b", xlab = "Date", ylab = "Cases", main = "Santa Clara Confirmed Cases - Timeseries")

```


## South Korea Summary

```{r, echo=FALSE}
sk_all = coronavirus %>% subset(Country.Region == "South Korea") 

sk_summary = sk_all %>% group_by(type) %>% summarise(cases = sum(cases))

mrate = sk_summary[2,2] / sk_summary[1,2]
mort_rate = data.frame("mort_rate", mrate)
names(mort_rate) = c("type", "cases")
sk_summary = rbind(sk_summary, mort_rate)

print(sk_summary)


```


## South Korea Trend

```{r, echo=FALSE}
sk_trend = sk_all %>% filter(type == "confirmed") %>% group_by(date)

plot(sk_trend$date, sk_trend$cases, type = "b", xlab = "Date", ylab = "Cases", main = "South Korea Confirmed Cases - Timeseries")

```

## Italy Summary

```{r, echo=FALSE}
it_all = coronavirus %>% subset(Country.Region == "Italy") 

it_summary = it_all %>% group_by(type) %>% summarise(cases = sum(cases))

mrate = it_summary[2,2] / it_summary[1,2]
mort_rate = data.frame("mort_rate", mrate)
names(mort_rate) = c("type", "cases")
it_summary = rbind(it_summary, mort_rate)

print(it_summary)


```


## Italy Trend

```{r, echo=FALSE}
it_trend = it_all %>% filter(type == "confirmed") %>% group_by(date)

plot(it_trend$date, it_trend$cases, type = "b", xlab = "Date", ylab = "Cases", main = "Italy Confirmed Cases - Timeseries")

```

## Iran Summary

```{r, echo=FALSE}
ir_all = coronavirus %>% subset(Country.Region == "Iran") 

ir_summary = ir_all %>% group_by(type) %>% summarise(cases = sum(cases))

mrate = ir_summary[2,2] / ir_summary[1,2]
mort_rate = data.frame("mort_rate", mrate)
names(mort_rate) = c("type", "cases")
ir_summary = rbind(ir_summary, mort_rate)

print(ir_summary)


```


## Iran Trend

```{r, echo=FALSE}
ir_trend = ir_all %>% filter(type == "confirmed") %>% group_by(date)

plot(ir_trend$date, ir_trend$cases, type = "b", xlab = "Date", ylab = "Cases", main = "Iran Confirmed Cases - Timeseries")

```


## France Summary

```{r, echo=FALSE}
fr_all = coronavirus %>% subset(Country.Region == "France") 

fr_summary = fr_all %>% group_by(type) %>% summarise(cases = sum(cases))

mrate = fr_summary[2,2] / fr_summary[1,2]
mort_rate = data.frame("mort_rate", mrate)
names(mort_rate) = c("type", "cases")
fr_summary = rbind(fr_summary, mort_rate)

print(fr_summary)


```


## France Trend

```{r, echo=FALSE}
fr_trend = fr_all %>% filter(type == "confirmed") %>% group_by(date)

plot(fr_trend$date, fr_trend$cases, type = "b", xlab = "Date", ylab = "Cases", main = "France Confirmed Cases - Timeseries")

```


## Japan Summary

```{r, echo=FALSE}
jp_all = coronavirus %>% subset(Country.Region == "Japan") 

jp_summary = jp_all %>% group_by(type) %>% summarise(cases = sum(cases))

mrate = jp_summary[2,2] / jp_summary[1,2]
mort_rate = data.frame("mort_rate", mrate)
names(mort_rate) = c("type", "cases")
jp_summary = rbind(jp_summary, mort_rate)

print(jp_summary)


```


## Japan Trend

```{r, echo=FALSE}
jp_trend = jp_all %>% filter(type == "confirmed") %>% group_by(date)

plot(jp_trend$date, jp_trend$cases, type = "b", xlab = "Date", ylab = "Cases", main = "Japan Confirmed Cases - Timeseries")

```


## Germany Summary

```{r, echo=FALSE}
gr_all = coronavirus %>% subset(Country.Region == "Germany") 

gr_summary = gr_all %>% group_by(type) %>% summarise(cases = sum(cases))

mrate = gr_summary[2,2] / gr_summary[1,2]
mort_rate = data.frame("mort_rate", mrate)
names(mort_rate) = c("type", "cases")
jp_summary = rbind(gr_summary, mort_rate)

print(gr_summary)


```


## Germany Trend

```{r, echo=FALSE}
gr_trend = gr_all %>% filter(type == "confirmed") %>% group_by(date)

plot(gr_trend$date, gr_trend$cases, type = "b", xlab = "Date", ylab = "Cases", main = "Germany Confirmed Cases - Timeseries")

```


## Spain Summary

```{r, echo=FALSE}
sn_all = coronavirus %>% subset(Country.Region == "Spain") 

sn_summary = sn_all %>% group_by(type) %>% summarise(cases = sum(cases))

mrate = sn_summary[2,2] / sn_summary[1,2]
mort_rate = data.frame("mort_rate", mrate)
names(mort_rate) = c("type", "cases")
sn_summary = rbind(sn_summary, mort_rate)

print(sn_summary)


```


## Spain Trend

```{r, echo=FALSE}
sn_trend = sn_all %>% filter(type == "confirmed") %>% group_by(date)

plot(sn_trend$date, sn_trend$cases, type = "b", xlab = "Date", ylab = "Cases", main = "Spain Confirmed Cases - Timeseries")

```